---
title: "final-project"
author: "Shaina Trevino"
date: "5/16/2019"
output: 
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(rio)
library(tidyverse)
library(broom)
```

## Importing data

```{r import}
drug_by_year <- import(here("data", "drug_by_year.csv"),
                       setclass = "tbl_df")

recent_drug <- import(here("data", "2015_drug_dep.csv"),
                      setclass = "tbl_df")
```

#### Info about the data:

`drug_by_year` has data from 2005 to 2015 and reports on the people who responded yes to using illicit substances in the last year. 

`recent_drug` has data from 2015-2016 (most recent NHANES data available) and shows the age of first drug use, a summed depression score, whether or not participants have been to rehab, have access to mental health services, and have health insurance. 
* `age_stim` is a composite measure of age of first meth and cocaine use (e.g., age of first stimulant use)

# Tutorial

## Introduction

What is looping/iterating

What is Functional Programming

## How to use `purrr::map`!

#### What is `purrr::map`

`purrr:map()` is a functional. It takes a vector (or list of vectors) and a function as it's input, runs the function for each element in the vector, and then returns a list.

  * The `purrr` package is loaded when you load `tidyverse`
  
For example, suppose you have a list that contains two other numbered lists with 5 values rounded to two decimal places and you wanted to round them all to the nearest integer.

Our list looks like this:

```{r create-list}
#creating list from two vectors of decimal values
v <- c(1.85, 2.07, 3.71, 4.21, 5.36)
v1 <- c(1.23, 2.53, 3.44, 4.71, 5.96)
l <- list(v, v1)
l
```

If we only had 1 vector to round, we could use the `round()` function

```{r ex-round}
round(v)
```

However, this does not work when supplied our list

```{r error-code, eval = FALSE}
round(l)
#FIGURE OUT HOW TO KEEP CODE AND ERROR WHEN KNIT FILE, BUT KNIT DOESNT WORK
```

Instead we need a function that will apply `round()` to every element in our list. Thats where we can use `map()`

```{r map-round}
#takes vector and rounds up
map(l, round)
```

Notice how the output from `map()` is a list, and the `round()` function has been applied to each element in the list. 

You could also use `map()` with a single vector, however in this case the output is the same as when solely using the `round()` function. 

```{r round-single-vec}
map(v, round)
```

However, as mentioned above, this output is a list. When you have an atomic vector as an input or a list of numbered lists of length 1, you can coerce your output to an atomic vector.

#### `map()` variants
  
You can override the return of list and instead return an atomic vector by using variants of `map()`:

    * `map_dbl` returns vector of type double
    
    * `map_chr` returns a vector of type character
    
    * `map_lgl` returns a vector of type logical (e.g., `TRUE` or `FALSE`)
    
    * `map_int` returns a vector of type integer
    
For example, if we use `map_dbl` or `map_chr` for our previous example, we will have the same results, but the output will be different types. 

```{r map-ex}
#return vector of type double
map_dbl(v, round)
#return vector as character
map_chr(v, round)
```

Notice how the numbers are actually the same, but `map_dbl` returns a numeric vector, whereas `map_chr` returns a character vector.

What if we wanted to know if each element in our vector is of type double, we would want to return a logical vector that tells us if the value is double (e.g., `TRUE`) or not (e.g., `FALSE`)

```{r map-ex1}
#return vector of type logical
map_lgl(v, is.double)
```

This logical vector tells us that each element is of type double.

`map_int` returns an integer vector (i.e., whole numbers), therefore, we can use it to ask questions like: what is the length of each element in my vector?

```{r map-ex2}
#return vector of type integer
map_int(v, length)
```

In this example, there is only 1 value per element in the vector, so all the lengths are 1.

If we use the same code on our list object, it should tell us how many numbers are in each element of the list.

```{r map-ex2-b}
map_int(l, length)
```

There are 2 elements in our list and each has 5 values. 

### Using `map()` in data frames

Since `map` can take a vector for an input, and data frames are lists containing vectors, you can use `map` to iterate through dataframes. 

EXPLAIN DATA

For example, say we want to use the NHANES data to calculate the mean for age of first use and depression scores. Since there is quite a bit of missing values in this data set, so we will need to add arguments to the mean function to ignore the missing data. 

Our data looks like this

```{r tidy-data-map-ex, echo = FALSE}
#mean and sd
d <- recent_drug %>% 
  select(age_cocaine:sum_depression)
d
```


To add arguments to a function within `map()`, we need to create an inline anonymous function where your function is first specified with `function(x)` where `x` is a place holder for what the function is going to run on. Then you write you function that you want to run on each element (`x`). 

Therefore, to calculate the means, our `map()` function would look like this

```{r summarize}
means <- map_dbl(d, function(x) mean(x, na.rm = TRUE)) 
means
```

There is also a shortcut you can use within `map()` functions where you don't have to specify `function(x)`. 

```{r summarize1}
means <- map_dbl(d, ~ mean(.x, na.rm = TRUE))
means
```

These two examples are exactly the same, the latter just has less code. `function(x)` is implied with the `~` and then you just have to use `.x` instead of `x` for what you are iterating over.

#### `map_df()`

There is another `map()` variant that is useful when working with dataframes, `map_df()`.

You can use `map_df()` to return data frame as the output instead of an atomic vector. 

For example, say we want to return data frame of the means and standard deviations of each column within our data frame. 

```{r map-df}
#you would expect this to work, but doesn't because only the last line in the function is being evaluated and printed
map_df(d, function(x) {
  mean(x, na.rm = TRUE)
  sd(x, na.rm = TRUE)
}) #note curly braces for longer functions

column_names <- colnames(d)
#need to instead
map_df(d, function(x) {
  summarize(d, mean = mean(x, na.rm = TRUE),
            sd = sd(x, na.rm = TRUE))
}, .id = "column_names")

#or you could do this in two separate steps, using map for looping the functions through vectors then putting vectors into dataframe
mean <- map(d, ~mean(.x, na.rm = TRUE))
sd <- map(d, ~sd(.x, na.rm = TRUE))
#combine vectors
cv <- c(mean, sd, recursive = TRUE)
#change attributes
attr(cv, "dim") <- c(6, 2) #6 variables in two columns (mean and sd)
colnames(cv) <- c("mean", "sd")
#transform to dataframe
cv <- data.frame(cv, row.names = column_names)
```

  
#### * At least one instance of parallel iteration (e.g., `map2_*`, `pmap_*`) 

EXPLAIN BOTH MAP 2 AND PMAP AND SHOW HOW PMAP WORKS WITH PLOT (BEFORE WALK? - SEE MIDTERM)

The `map` variants discussed above all take a single input to iterate (i.e., call the function on), but sometimes you need to iterate through more than just one input. 

This can be accomplished with `map2_` or `pmap_`

* `map2_` iterates over two inputs (e.g., two vectors)

* `pmap_` iterates over any number of inputs

For example, say you wanted to calculate a weighted mean from a vector of data values and a vector of weights.


```{r parallel}
data <- d %>% filter(!is.na(age_marijuana))
obs <- map(1:1715, ~rnorm(n = 20, mean = 17.289, sd = 4.558)) #1715 obs for marijuana, mean and sd from table above
wts <- map(1:1715, ~rpois(20, 5))

#this doesnt work because the weight argument is expecting a single value to apply to all observations. 
#map_dbl(obs, ~weighted.mean(.x, w = wts)) - had to comment to knit

#Instead we need a parallel iteration to run the function on each vector of observation with each corresponding vector of weights. 
map2_dbl(obs, wts, ~weighted.mean(.x, .y))
#make note of .x and .y and explain
#we now have a weighted mean score for each participant
```

`pmap()` can be supplied any number of inputs that you provide in a single list. This list can contain any number of inputs or arguments. 

`pmap()` is very useful when working with dataframes in which there are many vectors of the same length that can be provided as inputs (i.e., columns)

You can name the inputs in the list which is helpful when iterating through complex functions

*We will show you how to use `pmap` to iterate and build custom plots in a later blog post*


#### * At least one {purrr} function outside the basic `map` family (`walk_*`, `reduce`, `modify_*`, etc.) 

The `map()` variants discussed above all return an output (either an atomic vector or list). However, sometimes you don't need to return an atomic vector or list, but iteration would still be helpful.
  
Very useful variant of `map()` is `walk()`, specifically `walk2()` which can be used to save objects to a disk. Note: when saving something to a disk you have to provide 2 inputs: the object and the file path - this requires parallel iteration.   

```{r walk-ex, eval = FALSE}
by_year <- split(drug_by_year, drug_by_year$year)
paths <- here("data", paste0("year-", names(by_year), ".csv"))
walk2(by_year, paths, write.csv)
#splits our longitudinal dataset by year and saves separate csv file for each year with the title ("year-2005") in the data folder of the working directory

list.files("data")
```

`reduce()` takes a vector of any length as it's input and returns a vector of length 1. For example `reduce(1:4, f)` returns `f(f(f(1, 2), 3), 4)`. 

`modify()` returns the same type of input it is provided. For example, `modify(df, f)` will run the function on each column of the dataframe and return a dataframe as the output.


#### * At least one use case of `purrr::nest %>% mutate()` 

EXPLAIN LIST COLUMNS AND USE FOR GROUPING STATS (e.g., group by drug and year then calculate age of use etc.)

```{r nest-mutate1}
dat <- recent_drug %>% 
  select(-1:-2, -age_stim) %>% 
  pivot_longer(cols = age_cocaine:age_marijuana, names_to = c("dis", "drug"), names_sep = "_") %>% 
  select(-dis) %>% 
  rename(age_use = value) %>% 
  filter(!is.na(age_use)) %>% 
  mutate(mental_health = ifelse(mental_health == "Yes", "Yes", "No"),
         sex = fct_relevel(sex, ref = "Male"),
         mental_health = fct_relevel(mental_health, ref = "No"))

#fit one model per drug and see coefficients
by_drug_coefs <- dat %>% 
  group_by(drug) %>% 
  nest() %>% 
  mutate(model = map(data, ~lm(age_use ~ sex, data = .x)),
         coef = map(model, ~tidy(.x))) %>% 
  unnest(coef)

#Look at model fit stats per drug model - will be bad
by_drug_fit <- dat %>% 
  group_by(drug) %>% 
  nest() %>% 
  mutate(model = map(data, ~lm(age_use ~ sex, data = .x)),
         fit = map(model, ~glance(.x))) %>% 
  unnest(fit)

#mean age of use for males and females in first model
by_drug_sex <- dat %>% 
  group_by(drug) %>% 
  nest() %>% 
  mutate(model = map(data, ~lm(age_use ~ sex, data = .x)),
         coefs = map(model, coef),
         males = map_dbl(coefs, 1),
         females = males + map_dbl(coefs, 2)) %>% 
  select(drug, males, females)

#fit multiple models by year
by_drug_models <- dat %>% 
  group_by(drug) %>% 
  nest() %>% 
  mutate(m1 = map(data, ~lm(age_use ~ sex, data = .x)),
         m2 = map(data, ~lm(age_use ~ sex + mental_health, data = .x)),
         m3 = map(data, ~lm(age_use ~ sex * mental_health, data = .x)))


```


```{r nest-mutate, eval = FALSE, include = FALSE}
d_long <- drug_by_year %>% 
  select(-V1, -ethnicity) %>% 
  mutate(year = factor(year),
         drug = factor(drug),
         sex = factor(sex),
         response = factor(response))

by_year_drug <- d_long %>% 
  group_by(year, drug, count) %>% 
  nest() %>% 
  mutate(count = map(data, ~mutate(.x, count = n())))
```


Function2: Summarize Mean, SD, Min, and Max using by a Group_By Variabe

Now let's take the age function we have just created (evaluating whether age of first marijuana use was before or after the age of 18) and use it as a grouping variable to examine the mean, standard deviation, and min/max values for the age of first use variable in the "recent_drug,csv" data set. 
1) first we name the data set, the outcome variable, and the grouping variable. We will use these to call up this info when applying the completed function.
2) Next we use "enquo", part of the 'dplyr' package which should be loaded if you have not already done so, to select the variables we are planning to use in defining the function. Enquo helps R search for these variables within our data frame instead of searching the global environment for them. When using enquo the "!!" symbol will be neccessary anytime we want to refer back to these variables that we are defining here in the first steps of creating our function. EX. instead of writting mean(our_var_name) as usual we will need to write mean(!!our_var_name).
3) The %>% symbol is a pipe operator and is used to feed each piece of our function through the proceeding statement, so that at each step the result of the previous argument is fed through the following argument. Ex. dataset %>% step1 %>% step2.
4)"group_by" tells R to divide the outcome into a serperate row for each category of the grouping variable. 
5) "summarize" is the step where we envoke the operations and calculations on our outcome variable. The results of the summarize step should be seperated thanks to the group_by statement in step 4, 

The function can now be stored and run on our data. We can examine the mean first age of use by any of the other variables in our dataset, including the 18 or below variable we created in the first function we created. Other interesting grouping variables might be income, ethnicity, or mental health status. If you decide you want to group our outcome first by two variables this is explored in function 3.



```{r eval=FALSE}
library(dplyr)

fun2_mean1group <- function(data, cat_var, num_var){
  cat_var <- enquo(cat_var)
  num_var <- enquo(num_var)
  data %>% #if (cat_var2 < 17)
    group_by(!!cat_var) %>%
    summarize(mean = mean(!!num_var, na.rm = TRUE), sd = sd(!!num_var, na.rm = TRUE), min = min(!!num_var, na.rm = TRUE), max = max(!!num_var, na.rm = TRUE))
}

fun2_mean1group(recent_drug, mental_health, age_marijuana)
#fun2_mean1group(iris, Species, Sepal.Width)
```

Function 3: Two grouping variables 
5) If you wish to add a second "group_by" variable start by including a variable name in the Function itself, and include an additional enquo command
6) Now in the "group_by" command include a second variable using the "!!" as before. ex: group_by(!!group_var1, !!group_var2)
7) after the "summarize" command include an additional %>% and add the "spread" command naming the varable as before (!!group_var1), and indicating the outcome you wish to report. 


```{r}
fun3_mean2groups <- function(x, cat_var1, cat_var2, outcome_var){
  cat_var1 <- enquo(cat_var1)
  cat_var2 <- enquo(cat_var2)
  num_var <- enquo(outcome_var)
  
  x %>%
    group_by(!!cat_var1, !!cat_var2) %>%
    summarize(mean = mean(!!num_var, na.rm = TRUE),
              sd = sd(!!num_var, na.rm = TRUE), 
              min = min(!!num_var, na.rm = TRUE), 
              max = max(!!num_var, na.rm = TRUE))%>%
    spread(!!cat_var1, mean)
}

fun3_mean2groups(recent_drug, mental_health, income, age_marijuana)
#fun3_mean2groups(iris, Species, Petal.Width, Sepal.Width)

```

Function 4: Creating a Function Using GGPlot

To plot our outcome variable, in this case age of first marijuana use, we can create a function which will plot the age of first grouped by variable of interst. To test our function we will use "income" as category to group age of first marijuana use against on the x-axis. 

1) Once again we begin by naming the variables our function will use to pull data from. In this case we will include a data variable to name the dataset, "x" for our x-axis variable, and "y" for our y-axis variable. 
2) Next in the open bracket we will include our plotting function in this case ggplot, and we will name the dataset and x/y variables so they match the variable names we will use in the final function. 
3) Now we add the method we will use to plot, choosing one of the ggplot options. In this case we selected 'geom_point' which will provide a number of dots along the y-axis for each category. A color can also be selected. Ex. geom_point(color = "magenta"). 


```{r eval=FALSE}
fun4_plot <- function(data, x, y) {
    ggplot(data, aes_string(x, y)) +
      geom_point(color = "blue")
}

fun4_plot(recent_drug, "income", "age_marijuana")

```

Fununction 5: Plotting Mean, SD, Min, & Max

To plot the summary data for our age of first use variables (age_marijuan, age_meth, age_heroin, and age_cocaine) found in our "recent_drug.csv" data we can create a function which will plot all of our summary variables (mean, sd, min, and max) at once.

```{r eval=FALSE}
#not working can delete
  fun5_plotsummary <- function(filename) {
  dat <- read.csv(file = filename, header = FALSE)
  mean_age <- apply(dat, 2, mean, na.rm = TRUE)
  plot(mean_age)
  sd_age <- apply(dat, 2, sd, na.rm = TRUE)
  max_age <- apply(dat, 2, max, na.rm = TRUE)
  plot(max_age)
  min_age <- apply(dat, 2, min, na.rm = TRUE)
  plot(min_age)
}

fun4_plot(recent_drug)
```



## Requirements per syllabus

* *No code is used repetitively (no more than twice)*
* More than one variant of `purrr::map` is used 
* At least one {purrr} function outside the basic `map` family (`walk_*`, `reduce`, `modify_*`, etc.) 
* At least one instance of parallel iteration (e.g., `map2_*`, `pmap_*`) 
* At least one use case of `purrr::nest %>% mutate()` 

* At least two custom functions 
	+ Each function must be "pure" unless it is clear that it should have side 
	  effects (e.g., a plotting function)
  + Each function must do exactly one thing
  + The functions **may** replicate the behavior of a base function - as noted 
    above this is about practicing the skills you learn in class
* Code is fully reproducible and housed on GitHub 
* No obvious errors in chosen output format 
* Deployed on the web and shareable through a link 

