---
title: "final-project"
author: "Shaina Trevino"
date: "5/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(rio)
```

## Importing data

```{r import}
drug_by_year <- import(here("data", "drug_by_year.csv"),
                       setclass = "tbl_df")

recent_drug <- import(here("data", "2015_drug_dep.csv"),
                      setclass = "tbl_df")
```

#### Info about the data:

`drug_by_year` has data from 2005 to 2015 and reports on the people who responded yes to using illicit substances in the last year. 

`recent_drug` has data from 2015-2016 (most recent NHANES data available) and shows the age of first drug use, a summed depression score, whether or not participants have been to rehab, have access to mental health services, and have health insurance. 
  * `age_stim` is a composite measure of age of first meth and cocaine use (e.g., age of first stimulant use)

# Tutorial

## How to use `purrr::map`!

* More than one variant of `purrr::map` is used 

LIST ALL VARIANTS OF PURRR:MAP AND EXPLAIN DIFFERENCES - SHOW 2 TYPES

* At least one {purrr} function outside the basic `map` family (`walk_*`, 
  `reduce`, `modify_*`, etc.) 
  
EXPLAIN ALL (MOSTLY WALK) AND SHOW HOW IT IS USED - FOR PLOT EXAMPLE - SEE MIDTERM
  
* At least one instance of parallel iteration (e.g., `map2_*`, `pmap_*`) 

EXPLAIN BOTH MAP 2 AND PMAP AND SHOW HOW PMAP WORKS WITH PLOT (BEFORE WALK? - SEE MIDTERM)

* At least one use case of `purrr::nest %>% mutate()` 

EXPLAIN LIST COLUMNS AND USE FOR GROUPING STATS (e.g., group by drug and year then calculate age of use etc.) - POSSIBLY ON MIDTERM OR MACHINE LEARNING


## Requirements per syllabus

* *No code is used repetitively (no more than twice)*
* More than one variant of `purrr::map` is used 
* At least one {purrr} function outside the basic `map` family (`walk_*`, 
  `reduce`, `modify_*`, etc.) 
* At least one instance of parallel iteration (e.g., `map2_*`, `pmap_*`) 
* At least one use case of `purrr::nest %>% mutate()` 
* At least two custom functions 
	+ Each function must be "pure" unless it is clear that it should have side 
	  effects (e.g., a plotting function)
  + Each function must do exactly one thing
  + The functions **may** replicate the behavior of a base function - as noted 
    above this is about practicing the skills you learn in class
* Code is fully reproducible and housed on GitHub 
* No obvious errors in chosen output format 
* Deployed on the web and shareable through a link 

